---
title: "Coursera Practical Machine Learning"
author: "Zahra Fitriana"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
library(caret)
library(randomForest)
library(rpart)
library(ggplot2)
library(dplyr)

```{r data-prep}
# Load data
train_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

train_data <- read.csv(train_url, na.strings = c("NA", "", "#DIV/0!"))
test_data <- read.csv(test_url, na.strings = c("NA", "", "#DIV/0!"))

# Remove near-zero variance predictors
nzv <- nearZeroVar(train_data)
train_data <- train_data[, -nzv]

# Remove columns with too many NAs (keep columns with no NAs)
train_data <- train_data[, colSums(is.na(train_data)) == 0]

# Remove ID and timestamps columns (first 5)
train_data <- train_data[, -(1:5)]

# Make sure 'classe' is a factor
train_data$classe <- factor(train_data$classe)

# Apply some cleaning steps to test_data
test_data <- test_data[, names(test_data) %in% names(train_data)]
# test_data should not contain 'classe' column
test_data <- test_data[, names(train_data)[names(train_data) != "classe"]]

# Check final columns
print(paste("Number of columns in training data", ncol(train_data)))
print(paste("Number of columns in test data", ncol(test_data)))

```{r train_model}
set.seed(123)

# Split data into training and validation sets
inTrain <- createDataPartition(train_data$classe, p=0.7, list = FALSE)
train_set <- train_data[inTrain, ]
validation_set <- train_data[-inTrain, ]

# Train Random Forest model with 5-fold CV
control <- trainControl(method = "cv", number = 5)
rf_model <- train(classe ~ ., data = train_set, method = "rf", trControl = control)

print(rf_model)

```{r predict-evaluate}
# Predict on validation set
predictions <- predict(rf_model, validation_set)

# Make sure factors have the same levels
levels_all <- union(levels(predictions), levels(validation_set$classe))
predictions <- factor(predictions, levels = levels_all)
validation_set$classe <- factor(validation_set$classe, levels = levels_all)

# Confusion matrix
cm <- confusionMatrix(predictions, validation_set$classe)
print(cm)

# Compute and print error rate
error_rate <- 1 - cm$overall["Accuracy"]
print(paste("Validation error rate:", round(error_rate, 4)))

```{r final-predictions}
# Predict on test data (no classe column
final_predictions <- predict(rf_model, test_data)

print(final_predictions)

```{r save-predictions, echo=TRUE}
# Save 20 predictions into separate text files for Coursera quiz

pml_write_files <- function(x){
  n <- length(x)
  for(i in 1:n){
    filename <- paste0("problem_id_", i, ".txt")
    write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, col.names = FALSE)}
}

pml_write_files(final_predictions)